### 2017-12-14 submission/manuscript.Rmd
add paragraph on bray-curtis to discussion

***Statistical Analysis:***
All analysis was done with the R (v `r paste(version$major, ".", version$minor, sep = "")`) software package [@r_citation_2017]. Data tranformation and graphing was completed using the tidyverse package (v `r paste(packageVersion("tidyverse")[1])`) and colors selected using the viridis package (v `r paste(packageVersion("viridis")[1])`) [@tidyverse_2017; @viridis_citation_2017]. The total number of OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the fecal samples the data was normalized to each individual by cycle number to account for the biological variation between different people. Bray-Curtis distance matrices were generated using mothur after 100 sub-samplings at 1000, 5000, 10000, and 15000. The distance matrix data was analyzed using PERMANOVA with the vegan package (v `r paste(packageVersion("vegan")[1])`) [@vegan_citation] and kruskal-wallis tests within R. For both error and chimera analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc test. Where applicable correction for multiple comparison utilized the Benjamini-Hochberg method [@benjamini_controlling_1995].

***Analysis Workflow:***
The total number of OTUs after sub-sampling was analyzed for both the fecal and mock community samples. Cycle dependent affects on Bray-Curits distances were next assessed for the fecal samples looking at both overall cycle differences and within individual differences for the previous cycle (e.g. 20 versus 25, 25 versus 30, etc.). From these observations we wanted to next analyze potential reasons as to why some of these differences may have occured. First, analysis of general sequence error rate, number of sequences with an error, and base subsitution were assessed in the mock community for each Taq. After assessing these errors, the total number of chimeras was determined after sequence processing. The fecal samples were analyzed at 4 different sub-sampling levels, 1000, 5000, 10000, and 15000 while the mock community samples were analysed at 3 levels, 1000, 5000, 10000.

@@

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
Using PERMANOVA to test whether there were any differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Interestingly, Phusion was one of only two DNA polymerase that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5 cycle increments within each sample. We found that there was no detectable difference in distance when comparing to the previous 5 cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curits distance disappear for both differences in cycle number and within 5 cycle increments within an individual.

@@

>>>>

***Statistical Analysis:***
All analysis was done with the R (v `r paste(version$major, ".", version$minor, sep = "")`) software package [@r_citation_2017]. Data tranformation and graphing was completed using the tidyverse package (v `r paste(packageVersion("tidyverse")[1])`) and colors selected using the viridis package (v `r paste(packageVersion("viridis")[1])`) [@tidyverse_2017; @viridis_citation_2017]. The total number of OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the fecal samples the data was normalized to each individual by cycle number to account for the biological variation between different people. Bray-Curtis matrices were generated using mothur after 100 sub-samplings at 1000, 5000, 10000, and 15000. The distance matrix data was analyzed using PERMANOVA with the vegan package (v `r paste(packageVersion("vegan")[1])`) [@vegan_citation] and kruskal-wallis tests within R. For both error and chimera analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc test. Where applicable correction for multiple comparison utilized the Benjamini-Hochberg method [@benjamini_controlling_1995].

***Analysis Workflow:***
The total number of OTUs after sub-sampling was analyzed for both the fecal and mock community samples. Cycle dependent affects on Bray-Curits indicies were next assessed for the fecal samples looking at both overall cycle differences and within individual differences for the previous cycle (e.g. 20 versus 25, 25 versus 30, etc.). From these observations we wanted to next analyze potential reasons as to why some of these differences may have occured. First, analysis of general sequence error rate, number of sequences with an error, and base subsitution were assessed in the mock community for each Taq. After assessing these errors, the total number of chimeras was determined after sequence processing. The fecal samples were analyzed at 4 different sub-sampling levels, 1000, 5000, 10000, and 15000 while the mock community samples were analysed at 3 levels, 1000, 5000, 10000.

@@

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
Using PERMANOVA to test whether there were any differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Interestingly, Phusion was one of only two DNA polymerase that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5 cycle increments within each sample. We found that there was no detectable difference in Bray-Curits index when comparing to the previous 5 cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curits indicies disappear for both differences in cycle number and within 5 cycle increments within an individual.

@@

Although we did not observe strong differences, based on cycle number, using the Bray-Curtis index the data suggests that there may be differences between 15 cycles and higher cycle numbers, such as 30, that are commonly used. What is most interesting is that there was no difference within individuals between corresponding 5 cycle increments (e.g. 15 to 20, 20 to 25, etc.). This finding, in conjunction with the PERMANOVA results, would suggest that differences based on Bray-Curtis incrementaly build up so that within 5 cycle increments the differences are not that large but when comparing larger cycle differences, as with the PERMANOVA, these differences start to become detectable. Although our data suggests that it is possible to overcome these types of cycle dependent differences with increased sub-sampling depth, for some DNA polymerase increased sub-sampling depth increases chimera prevalence and per base error rate. Conversely, it is possible that many of the increased number of OTUs, generated as cycle number increases, are not highly abundant and the Bray-Curtis index is able to successfully downweight these respective OTUs [@bc_index_Minchin1987].
